---
title: "Emotions in social media"
author: "Marc Cervera Rosell"
date: "2024-08-15"
output: pdf_document
---

# 1. Load dataset

[\underline{Dataset link}](https://www.kaggle.com/code/saadatkhalid/social-media-vs-emotions-eda-model-99-acc/input?select=train.csv>)

```{r}
tryCatch({
  data <- read.csv("train.csv", header = TRUE)
  print("File read successfully")
}, error = function(e) {
  cat("ERROR when loading the dataset",conditionMessage(e), "\n")
})
```


# 2. Preprocessing data

## 2.1 Delete blank lines (if needed):

```{r}
cat("Rows before:", nrow(data), "\n")
data <- data[rowSums(is.na(data) | data == "") != ncol(data), ]
cat("Rows after:", nrow(data))
```

## 2.2 Check variable types and column names

```{r}
columns <- names(data)
types <- sapply(data, class)
for (i in seq_along(columns)) {
  cat("Column name:", columns[i], " Type:", types[i], "\n")
}
```

Transformations:

- Column "Age" will become an integer

- Column "Daily_Usage_Time..minutes" will be renamed as "Minutes_Per_Day"

```{r}
data_transformed <- transform(data,
                              Age = as.integer(Age))
colnames(data_transformed)[colnames(data_transformed) ==
                             "Daily_Usage_Time..minutes."] <- "Minutes_Per_Day"
```

```{r}
types <- sapply(data_transformed, class)
for (i in seq_along(columns)) {
  cat("Column name:", columns[i], " Type:", types[i], "\n")
}
```

## 2.3 Check if there's NA values

```{r}
any(is.na(data_transformed))
```


# 3. Descriptive analysis and inferential

## 3.1 Data distribution per gender, platform, age and dominant emotion

### 3.1.1 Data distribution per gender

```{r}
genders <- unique(data_transformed$Gender)
print(genders)
```

```{r}
percentages_gender <- round(prop.table(table(data_transformed$Gender)) * 100, 2)
for (i in seq_along(genders)) {
  cat("Gender:",  genders[i], "- Percentage:", percentages_gender[i],"%\n")
}
```

```{r}
pie(table(data_transformed$Gender), main = "Distribution per age",
    col = rainbow(length(unique(data_transformed$Gender))))
    labels = genders
```

### 3.1.2 Data distribution per platform

```{r}
platforms <- unique(data_transformed$Platform)
print(platforms)
```

```{r}
percentages_platform <- round(prop.table(table(data_transformed$Platform)) * 100, 2)
for (i in seq_along(platforms)) {
  cat("Gender:",  platforms[i], "- Percentage:", percentages_platform[i],"%\n")
}
```

```{r}
pie(table(data_transformed$Platform), main = "Distribution per platform",
    col = rainbow(length(unique(data_transformed$Platform))))
    labels = platforms
```

### 3.1.3 Data distribution per age

```{r}
ages <- unique(data_transformed$Age)
print(ages)
```

```{r}
percentages_ages <- round(prop.table(table(data_transformed$Age)) * 100,
                          2)
for (i in seq_along(ages)) {
  cat("Gender:",  ages[i], "- Percentage:", percentages_ages[i],"%\n")
}
```

```{r}
pie(table(data_transformed$Age), main = "Distribution per age",
    col = rainbow(length(unique(data_transformed$Age))))
    labels = ages
```

### 3.1.4 Data distribution per dominant emotion

```{r}
emotions <- unique(data_transformed$Dominant_Emotion)
print(emotions)
```

```{r}
percentages_emotions <- round(prop.table(table(data_transformed$Dominant_Emotion)) * 100, 2)
for (i in seq_along(emotions)) {
  cat("Gender:",  emotions[i], "- Percentage:", percentages_emotions[i],"%\n")
}
```

```{r}
pie(table(data_transformed$Dominant_Emotion), main = "Distribution per dominant emotion",
    col = rainbow(length(unique(data_transformed$Dominant_Emotion))))
    labels = emotions
```

## 3.2 Check if there are extreme ages

```{r}
boxplot(data_transformed$Age, main = "Boxplot of ages",
        ylab = "Ages")
```

```{r}
cat("Median: ",median(data_transformed$Age),"\n")
cat("Quantiles 25%, 50%, 75%: ",quantile(data_transformed$Age, probs = c(0.25, 0.5, 0.75)))
```

Median line -> Closer to Q3 (not too much) -> More people has less than 27 YO.

Box -> Q1 = 24 and Q3 = 30.25 ~ 30 -> 50% of the people has between 24 and 30 YO.

Whiskers -> Top whisker is longer than bottom whisker -> Ages above median are more dispersed (More difference between the median and the max value)

Outliers -> There are no outliers

## 3.3 Relation between used platform and dominant emotion

H0 -> There's no significant association between both variables

H1 -> There's significant association between both variables

```{r}
chisq_data <- table(data_transformed$Platform, data_transformed$Dominant_Emotion)
chisq_data
```

```{r}
# Alpha = 0.05 -> CL = 95%
chisq.test(chisq_data, correct = FALSE)
```

The p value < 0.05(alpha) -> There's enough evidence to refuse H0 with a 95% confidence level

## 3.4 Dominant emotion depending on the minutes per day spend on social media

H0 -> The spent time on social media is the same regardless of the dominant emotion

H1 -> At least one emotion spends more/less time on social media

```{r}
plot(density(data_transformed$Minutes_Per_Day),
     main = "Minutes dedicated to social media per day",
     xlab = "Minutes",
     ylab = "Density",
     col = "blue",
     lwd = 2)
values_normal_distribution <- seq(min(data_transformed$Minutes_Per_Day),
                                  max(data_transformed$Minutes_Per_Day),
                                  length = 100)
normal_distribution <- dnorm(values_normal_distribution,
                             mean = abs(mean(data_transformed$Minutes_Per_Day)),
                             sd = sd(data_transformed$Minutes_Per_Day))
lines(values_normal_distribution, normal_distribution, col = "red", lwd = 2)
```

Variable Minutes_Per_Day -> No normal distribution -> ANOVA no possible

```{r}
kruskal <- kruskal.test(Minutes_Per_Day ~ Dominant_Emotion, data = data_transformed)
kruskal
```

p < 0.05 -> There's enough evidence to refuse H0 with a 95% confidence level -> Suggests relation

```{r}
library(dunn.test)
```

```{r}
results_dunn <- dunn.test(data_transformed$Minutes_Per_Day,
                          data_transformed$Dominant_Emotion, method = "bonferroni")
print(results_dunn)
```

Significant difference = Between both emotions, one of them spends more/less time on social media than the other emotion.

Considering significance = 0.05 -> CL = 95%:
  - "Anger - Boredom"
  
  - "Anxiety - Boredom"
  
  - "Anger - Happiness" 
  
  
  - "Anxiety - Happiness"
  
  - "Boredom - Happiness"
  
  - "Anger - Neutral"
  
  - "Anxiety - Neutral"
  
  - "Boredom - Neutral"
  
  - "Happiness - Neutral"
  
  - "Anger - Sadness"

```{r}
library(ggplot2)
library(dplyr)
```


```{r}
ggplot(data_transformed, aes(x = Dominant_Emotion, y = Minutes_Per_Day)) +
  geom_violin() +
  scale_y_continuous(limits = c(0, 200),
                     breaks = seq(0, 200, by = 20)) +
  labs(title = "Distribution of the minutes dedicated to social media
       per dominant emotion",
       x = "Dominant emotion",
       y = "Minutes per day")
```

Taking a look at the graphic above, H1 is confirmed.

```{r}
minutes_platform <- data_transformed %>%
  group_by(Platform) %>%
  summarise(suma = sum(Minutes_Per_Day)) %>%
  arrange(desc(suma))
print(minutes_platform)
```

## 3.5 Dominant emotion depending on the posts per day spend on social media

H0 -> The daily posts are the same regardless of the dominant emotion

H1 -> At least one emotion has more/less daily posts

```{r}
plot(density(data_transformed$Posts_Per_Day),
     main = "Posts per day on social media per day",
     xlab = "Posts",
     ylab = "Density",
     col = "blue",
     lwd = 2)
values_normal_distribution <- seq(min(data_transformed$Posts_Per_Day),
                                  max(data_transformed$Posts_Per_Day),
                                  length = 100)
normal_distribution <- dnorm(values_normal_distribution,
                             mean = abs(mean(data_transformed$Posts_Per_Day)),
                             sd = sd(data_transformed$Posts_Per_Day))
lines(values_normal_distribution, normal_distribution, col = "red", lwd = 2)
```

Variable Posts_Per_Day -> No normal distribution -> ANOVA no possible

```{r}
kruskal <- kruskal.test(Posts_Per_Day ~ Dominant_Emotion, data = data_transformed)
kruskal
```

p < 0.05 -> There's enough evidence to refuse H0 with a 95% confidence level -> Suggests relation

```{r}
results_dunn <- dunn.test(data_transformed$Posts_Per_Day,
                          data_transformed$Dominant_Emotion, method = "bonferroni")
print(results_dunn)
```

Significant difference = Between both emotions, one of them posts more/less on social media than the other emotion.

Considering significance = 0.05 -> CL = 95%:
  - "Anger - Anxiety"
  
  - "Anger - Boredom"
  
  - "Anxiety - Boredom"
  
  - "Anger - Happiness"
  
  - "Anxiety - Happiness"
  
  - "Boredom - Happiness"
  
  - "Anger - Neutral"
  
  - "Anxiety - Neutral"
  
  - "Boredom - Neutral"
  
  - "Happiness - Neutral"
  
  - "Anger - Sadness" 
  
  - "Boredom - Sadness"
  
  - "Happiness - Sadness"
  
  - "Neutral - Sadness"  
  
```{r}
ggplot(data_transformed, aes(x = Dominant_Emotion, y = Posts_Per_Day)) +
  geom_violin() +
  scale_y_continuous(limits = c(0, 9),
                     breaks = seq(0, 9, by = 1)) +
  labs(
    title = "Distribution of the posts per day on social media
    per dominant emotion",
       x = "Dominant emotion",
       y = "Posts per day")
```
  
Taking a look at the graphic above, H1 is confirmed.

```{r}
posts_platform <- data_transformed %>%
  group_by(Platform) %>%
  summarise(suma = sum(Posts_Per_Day)) %>%
  arrange(desc(suma))
print(posts_platform)
```  
  
  
## 3.6 Dominant emotion depending on the likes received per day spend on social media

H0 -> The likes received are the same regardless of the dominant emotion

H1 -> At least one emotion has more/less likes

```{r}
plot(density(data_transformed$Likes_Received_Per_Day),
     main = "Likes per day on social media per day",
     xlab = "Likes",
     ylab = "Density",
     col = "blue",
     lwd = 2)
values_normal_distribution <- seq(min(data_transformed$Likes_Received_Per_Day),
                                  max(data_transformed$Likes_Received_Per_Day),
                                  length = 100)
normal_distribution <- dnorm(values_normal_distribution,
                             mean = abs(mean(data_transformed$Likes_Received_Per_Day)),
                             sd = sd(data_transformed$Likes_Received_Per_Day))
lines(values_normal_distribution, normal_distribution, col = "red", lwd = 2)
```
  
Variable Likes_Received_Per_Day -> No normal distribution -> ANOVA no possible

```{r}
kruskal <- kruskal.test(Likes_Received_Per_Day ~ Dominant_Emotion, data = data_transformed)
kruskal
```

p < 0.05 -> There's enough evidence to refuse H0 with a 95% confidence level -> Suggests relation

```{r}
ggplot(data_transformed, aes(x = Dominant_Emotion, y = Likes_Received_Per_Day)) +
  geom_violin() +
    scale_y_continuous(limits = c(0, 120),
                     breaks = seq(0, 120, by = 10)) +
  labs(
    title ="Distribution of the likes received per day on social media
    per dominant emotion",
       x = "Dominant emotion",
       y = "Likes per day")
```
  
Taking a look at the graphic above, H1 is confirmed.
 
```{r}
likes_platform <- data_transformed %>%
  group_by(Platform) %>%
  summarise(suma = sum(Likes_Received_Per_Day)) %>%
  arrange(desc(suma))
print(likes_platform)
```  
  
## 3.7 Dominant emotion depending on the comments received per day spend on social media

H0 -> The comments received are the same regardless of the dominant emotion

H1 -> At least one emotion has more/less comments

```{r}
plot(density(data_transformed$Comments_Received_Per_Day),
     main = "Comments per day on social media per day",
     xlab = "Comments",
     ylab = "Density",
     col = "blue",
     lwd = 2)
values_normal_distribution <- seq(min(data_transformed$Comments_Received_Per_Day),
                                  max(data_transformed$Comments_Received_Per_Day),
                                  length = 100)
normal_distribution <- dnorm(values_normal_distribution,
                             mean = abs(mean(data_transformed$Comments_Received_Per_Day)),
                             sd = sd(data_transformed$Comments_Received_Per_Day))
lines(values_normal_distribution, normal_distribution, col = "red", lwd = 2)
```
  
Variable Comments_Received_Per_Day -> No normal distribution -> ANOVA no possible

```{r}
kruskal <- kruskal.test(Comments_Received_Per_Day ~ Dominant_Emotion, data = data_transformed)
kruskal
```

p < 0.05 -> There's enough evidence to refuse H0 with a 95% confidence level -> Suggests relation

```{r}
results_dunn <- dunn.test(data_transformed$Comments_Received_Per_Day,
                          data_transformed$Dominant_Emotion, method = "bonferroni")
print(results_dunn)
```

Significant difference = Between both emotions, one of them posts more/less on social media than the other emotion.

Considering significance = 0.05 -> CL = 95%:

  - "Anger - Anxiety"
  
  - "Anger - Boredom"
  
  - "Anxiety - Boredom"
  
  - "Anger - Happiness"
  
  - "Anxiety - Happiness"
  
  - "Boredom - Happiness"
  
  - "Anger - Neutral"
  
  - "Anxiety - Neutral"
  
  - "Boredom - Neutral"
  
  - "Happiness - Neutral"
   
  - "Anger - Sadness"
   
  - "Anxiety - Sadness"  

  - "Happiness - Sadness"
  
  - "Neutral - Sadness"


```{r}
ggplot(data_transformed, aes(x = Dominant_Emotion, y = Comments_Received_Per_Day)) +
  geom_violin() +
    scale_y_continuous(limits = c(0, 45),
                     breaks = seq(0, 45, by = 10)) +
  labs(
    title = "Distribution of the comments received per day on social media per
    dominant emotion",
       x = "Dominant emotion",
       y = "Comments per day")
``` 
  
Taking a look at the graphic above, H1 is confirmed.

## 3.8 Dominant emotion depending on the messages sent per day spend on social media

H0 -> The messages sent are the same regardless of the dominant emotion

H1 -> At least one emotion has more/less messages

```{r}
plot(density(data_transformed$Messages_Sent_Per_Day),
     main = "Messages per day on social media per day",
     xlab = "Messages",
     ylab = "Density",
     col = "blue",
     lwd = 2)
values_normal_distribution <- seq(min(data_transformed$Messages_Sent_Per_Day),
                                  max(data_transformed$Messages_Sent_Per_Day),
                                  length = 100)
normal_distribution <- dnorm(values_normal_distribution,
                             mean = abs(mean(data_transformed$Messages_Sent_Per_Day)),
                             sd = sd(data_transformed$Messages_Sent_Per_Day))
lines(values_normal_distribution, normal_distribution, col = "red", lwd = 2)  
``` 
    
Variable Comments_Received_Per_Day -> No normal distribution -> ANOVA no possible

```{r}
kruskal <- kruskal.test(Messages_Sent_Per_Day ~ Dominant_Emotion, data = data_transformed)
kruskal
```

p < 0.05 -> There's enough evidence to refuse H0 with a 95% confidence level -> Suggests relation

```{r}
messages_dunn <- dunn.test(data_transformed$Messages_Sent_Per_Day,
                          data_transformed$Dominant_Emotion, method = "bonferroni")
print(results_dunn)
```

Significant difference = Between both emotions, one of them send more/less messages on social media than the other emotion.

Considering significance = 0.05 -> CL = 95%:

  - "Anger - Anxiety"
  
  - "Anger - Boredom"
  
  - "Anxiety - Boredom"
  
  - "Anger - Happiness"
  
  - "Anxiety - Happiness"
  
  - "Boredom - Happiness"
  
  - "Anger - Neutral" 
  
  - "Anxiety - Neutral"
  
  - "Boredom - Neutral"
  
  - "Happiness - Neutral"
  
  - "Anger - Sadness"
  
  - "Anxiety - Sadness"
  
  - "Boredom - Sadness"

```{r}
ggplot(data_transformed, aes(x = Dominant_Emotion, y = Messages_Sent_Per_Day)) +
  geom_violin() +
      scale_y_continuous(limits = c(0, 55),
                     breaks = seq(0, 55, by = 10)) +
  labs(
    title = "Distribution of the messages sent per day on social media per
    dominant emotion",
       x = "Dominant emotion",
       y = "Messages per day")
``` 

Taking a look at the graphic above, H1 is confirmed.

## 3.9 Relation between dominant emotion and gender

H0 -> There's no significant association between both variables

H1 -> There's significant association between both variables

```{r}
chisq_data_2 <- table(data_transformed$Gender, data_transformed$Dominant_Emotion)
chisq_data_2
```

```{r}
# Alpha = 0.05 -> CL = 95%
chisq.test(chisq_data_2, correct = FALSE)
```

The p value < 0.05(alpha) -> There's enough evidence to refuse H0 with a 95% confidence level

## 3.10 Dominant emotion depending on age

H0 -> The age is the same regardless of the dominant emotion

H1 -> At least one emotion has more/less age

```{r}
plot(density(data_transformed$Age),
     main = "Age on social media per day",
     xlab = "Age",
     ylab = "Density",
     col = "blue",
     lwd = 2,
     ylim = c(0, 0.10))
values_normal_distribution <- seq(min(data_transformed$Age),
                                  max(data_transformed$Age),
                                  length = 1000)
normal_distribution <- dnorm(values_normal_distribution,
                             mean = abs(mean(data_transformed$Age)),
                             sd = sd(data_transformed$Age))
lines(values_normal_distribution, normal_distribution, col = "red", lwd = 2)
``` 

Variable Age -> No normal distribution -> ANOVA no possible

```{r}
kruskal <- kruskal.test(Age ~ Dominant_Emotion, data = data_transformed)
kruskal
```

p < 0.05 -> There's enough evidence to refuse H0 with a 95% confidence level -> Suggests relation

```{r}
messages_dunn <- dunn.test(data_transformed$Age,
                          data_transformed$Dominant_Emotion, method = "bonferroni")
print(results_dunn)
```

Significant difference = Between both emotions, one of them has more/less age

Considering significance = 0.05 -> CL = 95%:

  - "Anger - Anxiety"
  
  - "Anger - Boredom"
  
  - "Anxiety - Boredom"
  
  - "Anger - Happiness"
  
  - "Anxiety - Happiness"
  
  - "Boredom - Happiness"
  
  - "Anger - Neutral"    
 
  - "Anxiety - Neutral"
  
  - "Boredom - Neutral"
  
  - "Happiness - Neutral"
  
  - "Anger - Sadness"

```{r}
ggplot(data_transformed, aes(x = Dominant_Emotion, y = Age)) +
  geom_violin() +
        scale_y_continuous(limits = c(20, 40),
                     breaks = seq(20, 40, by = 5)) +
  labs(
    title = "Distribution of the messages sent per day on social media per
    dominant emotion",
       x = "Dominant emotion",
       y = "Messages per day")
```

Taking a look at the graphic above, H1 is confirmed.

# 4. Random forest prediction

```{r}
library(randomForest)
library(caret)
```

```{r}
set.seed(123)
index <- sample(1:nrow(data_transformed), 0.8 * nrow(data_transformed))
training_data <- data_transformed[index, ]
test_data <- data_transformed[-index, ]
```

```{r}
control <- trainControl(method = "repeatedcv", number = 10, repeats = 3,
                        search = "random")
model_random_forest <- train(Dominant_Emotion ~ Gender + Age + Platform +
                                      Minutes_Per_Day + Posts_Per_Day + 
                                      Likes_Received_Per_Day + 
                                      Comments_Received_Per_Day + 
                                      Messages_Sent_Per_Day,
                             data = training_data,
                             method = "rf",
                             metric = "Accuracy",
                             tuneLength = 13,
                             trControl = control)
print(model_random_forest)
```

```{r}
predictions <- predict(model_random_forest, test_data)
confusionMatrix(as.factor(predictions), as.factor(test_data$Dominant_Emotion))
```

Test data accuracy is 97.5%

```{r}
plot(model_random_forest)
```